---
title: 'Assignment #4'
author: "Elliot Smith"
date: "9/17/2018"
output: pdf_document
---

# Problem 1

```{r, echo = FALSE}

## Load necessary packages

library(coda)
suppressMessages(
suppressWarnings(
    library(asbio)
)
)



########## Problem 1 ##########



##### Part i #####



num_draws <- 10^4

mean_c <- 1.013
mean_t <- 1.173

## Draw from the scaled inverse chi-square

n_1 <- 32
df <- n_1 - 1

samp_i_1 <- rinvchisq(num_draws, df, scale=1/df)

## Draw from the normal distribution

samp_i_2 <- rnorm(num_draws, mean = mean_c, sqrt(samp_i_1))



##### Part ii #####

## Draw from the scaled inverse chi-square

n_2 <- 36
df <- n_2 - 1

samp_ii_1 <- rinvchisq(num_draws, df, scale=1/df)

## Draw from the normal distribution

samp_ii_2 <- rnorm(num_draws, mean = mean_t, sqrt(samp_ii_1))

```

## Part i

Please see the Code Appendix for the sampling technique.

## Part ii

Please see the Code Appendix for the sampling technique.

## Part iii

```{r, echo = FALSE, fig.height = 4}

##### Part iii #####

## Generate the histogram of mean differences

hist(samp_ii_2 - samp_i_2,
     xlab = "Mean Difference",
     ylab = "",
     main ="Histogram: Mean Difference")

```

## Part iv

```{r, echo = FALSE}

##### Part iv #####

## Construct the HPD interval of mean differences

interval <- as.vector(HPDinterval(as.mcmc(samp_ii_2 - samp_ii_1), prob = 0.95)[1, 1:2])

## Output the result

cat("HPD Interval: ", "(", interval[1], ",", interval[2], ")")

```

Since out interval does not contain zero, we can be certain that the treatment group mean is larger than the control group mean, thus implying that the treatment does have a material affect as compared to the control.

# Problem 2

## Part i

Prior Distribution:
\newline
\[
\begin{aligned}
p(\theta) & \propto exp\bigg[-\frac{1}{2}\big(\theta - \mu_\theta\big)^{T}\Sigma_\theta^{-1}(\theta - \mu_\theta)\bigg] \\
& \propto exp\bigg[-\frac{1}{2}\big(\theta^{T}\Sigma_\theta^{-1}{\theta} - 2\theta^{T}\Sigma_\theta^{-1}\mu_\theta + \mu_\theta^{T}\Sigma_\theta^{-1}\mu_\theta\bigg] \\
& \propto exp\bigg[-\frac{1}{2}\big(\theta^{T}\Sigma_\theta^{-1}{\theta} - 2\theta^{T}\Sigma_\theta^{-1}\mu_\theta + \mu_\theta^{T}\Sigma_\theta^{-1}\mu_\theta\big)\bigg] \\
& \propto exp\bigg[-\frac{1}{2}\big(\theta^{T}\Sigma_\theta^{-1}{\theta} - 2\theta^{T}\Sigma_\theta^{-1}\mu_\theta\big)\bigg] \\
& \propto exp\bigg[-\frac{1}{2}\big(\theta^{T}Q_{\theta_1}{\theta} - 2\theta^{T}\ell_{\theta_1}\big)\bigg] \\
\end{aligned}
\]
\newline
Where:
\newline
\newline
$Q_{\theta_1} = \Sigma_\theta^{-1}$
\newline
$\ell_{\theta_1} = \Sigma_\theta^{-1}\mu_\theta$
\newline
\newline
\newline
Likelihood:
\newline
\[
\begin{aligned}
p(y|\theta) & \propto exp\bigg[-\frac{1}{2}\bigg(\sum_{i=1}^n y_i^{T}\Sigma^{-1}y_i - 2\sum_{i=1}^n \theta^T{\Sigma^{-1}}y_i + \sum_{i=1}^n \theta^T \Sigma^{-1}\theta \bigg)\bigg] \\
& \propto exp\bigg[-\frac{1}{2}\bigg(\theta^T n\Sigma^{-1}\theta - 2 \theta^T{\Sigma^{-1}}n{\bar{y}}\bigg)\bigg] \\
& \propto exp\bigg[-\frac{1}{2}\bigg(\theta^TQ_{\theta_2}\theta - 2 \theta^T\ell_{\theta_2}\bigg)\bigg] \\
\end{aligned}
\]
\newline
Where:
\newline
\newline
$Q_{\theta_2} = n\Sigma^{-1}$
\newline
$\ell_{\theta_2} = \Sigma^{-1}n\bar{y}$
\newline
\newline
\newline
Posterior:
\newline
\[
\begin{aligned}
p(\theta|y) & \propto exp\bigg[-\frac{1}{2}\big(\theta^{T}Q_{\theta_1}{\theta} - 2\theta^{T}\ell_{\theta_1}\big)\bigg] \times
exp\bigg[-\frac{1}{2}\bigg(\theta^TQ_{\theta_2}\theta - 2 \theta^T\ell_{\theta_2}\bigg)\bigg] \\
& \propto exp\bigg[-\frac{1}{2}\theta^TQ_{\theta_1}\theta + \theta^T\ell_{\theta_1} -\frac{1}{2}\theta^TQ_{\theta_2}\theta + \theta^T\ell_{\theta_2}\bigg] \\
& \propto exp\bigg[\theta^T\big(\ell_{\theta_1} + \ell_{\theta_2}\big) - \frac{1}{2}\theta^T\big(Q_{\theta_1} + Q_{\theta_2}\big)\theta\bigg] \\
& \propto exp\bigg[-\frac{1}{2}\big(\theta^TQ_\theta\theta - 2\theta^T\ell_\theta\big)\bigg] \\
\end{aligned}
\]
\newline
Where:
\newline
\newline
$Q_{\theta} = Q_{\theta_1} + Q_{\theta_2} = \Sigma_\theta^{-1} + n\Sigma_\theta^{-1}$
\newline
$\ell_{\theta} = \ell_{\theta_1} + \ell_{\theta_2} = \Sigma_\theta^{-1}\mu_\theta + \Sigma^{-1}n\bar{y}$
\newline
\newline
Thus:
\newline
\newline
$\theta|y \sim MVN(Q_\theta^{-1}\ell_\theta, Q_\theta^{-1})$

## Part ii

Posterior Mean, $Q_\theta^{-1}\ell_\theta$:

$$
\begin{pmatrix} 
\frac{1}{\sigma_{\theta_1}^2} & 0 & 0 & 0 & ...\\ 
0 & \frac{1}{\sigma_{\theta_2}^2} & 0 & 0 & ...\\ 
0 & 0 & \frac{1}{\sigma_{\theta_3}^2} & 0 & ...\\ 
0 & 0 & 0 & \frac{1}{\sigma_{\theta_4}^2} & ... \\
\vdots & \vdots & \vdots & \vdots & \ddots
\end{pmatrix}^{-1}
+
\begin{pmatrix} 
\frac{n}{\sigma_{1}^2} & 0 & 0 & 0 & ...\\ 
0 & \frac{n}{\sigma_{2}^2} & 0 & 0 & ...\\ 
0 & 0 & \frac{n}{\sigma_{3}^2} & 0 & ...\\ 
0 & 0 & 0 & \frac{n}{\sigma_{4}^2} & ... \\
\vdots & \vdots & \vdots & \vdots & \ddots
\end{pmatrix}^{-1}
\times
\begin{pmatrix} 
\frac{1}{\sigma_{\theta_1}^2} & 0 & 0 & 0 & ...\\ 
0 & \frac{1}{\sigma_{\theta_2}^2} & 0 & 0 & ...\\ 
0 & 0 & \frac{1}{\sigma_{\theta_3}^2} & 0 & ...\\ 
0 & 0 & 0 & \frac{1}{\sigma_{\theta_4}^2} & ... \\
\vdots & \vdots & \vdots & \vdots & \ddots
\end{pmatrix}\mu_\theta
+
\begin{pmatrix} 
\frac{n\bar{y}}{\sigma_{1}^2} & 0 & 0 & 0 & ...\\ 
0 & \frac{n\bar{y}}{\sigma_{2}^2} & 0 & 0 & ...\\ 
0 & 0 & \frac{n\bar{y}}{\sigma_{3}^2} & 0 & ...\\ 
0 & 0 & 0 & \frac{n\bar{y}}{\sigma_{4}^2} & ... \\
\vdots & \vdots & \vdots & \vdots & \ddots
\end{pmatrix}
$$

Posterior Covariance, $Q_\theta^{-1}$:

$$
\begin{pmatrix} 
\frac{1}{\sigma_{\theta_1}^2} & 0 & 0 & 0 & ...\\ 
0 & \frac{1}{\sigma_{\theta_2}^2} & 0 & 0 & ...\\ 
0 & 0 & \frac{1}{\sigma_{\theta_3}^2} & 0 & ...\\ 
0 & 0 & 0 & \frac{1}{\sigma_{\theta_4}^2} & ... \\
\vdots & \vdots & \vdots & \vdots & \ddots
\end{pmatrix}^{-1}
+
\begin{pmatrix} 
\frac{n}{\sigma_{1}^2} & 0 & 0 & 0 & ...\\ 
0 & \frac{n}{\sigma_{2}^2} & 0 & 0 & ...\\ 
0 & 0 & \frac{n}{\sigma_{3}^2} & 0 & ...\\ 
0 & 0 & 0 & \frac{n}{\sigma_{4}^2} & ... \\
\vdots & \vdots & \vdots & \vdots & \ddots
\end{pmatrix}^{-1}
$$

## Part iii

Posterior Mean, $Q_\theta^{-1}\ell_\theta$:

$$
\begin{pmatrix} 
\frac{1}{\sigma_{\theta}^2} & 0 & 0 & 0 & ...\\ 
0 & \frac{1}{\sigma_{\theta}^2} & 0 & 0 & ...\\ 
0 & 0 & \frac{1}{\sigma_{\theta}^2} & 0 & ...\\ 
0 & 0 & 0 & \frac{1}{\sigma_{\theta}^2} & ... \\
\vdots & \vdots & \vdots & \vdots & \ddots
\end{pmatrix}^{-1}
+
\begin{pmatrix} 
\frac{n}{\sigma^2} & 0 & 0 & 0 & ...\\ 
0 & \frac{n}{\sigma^2} & 0 & 0 & ...\\ 
0 & 0 & \frac{n}{\sigma^2} & 0 & ...\\ 
0 & 0 & 0 & \frac{n}{\sigma^2} & ... \\
\vdots & \vdots & \vdots & \vdots & \ddots
\end{pmatrix}^{-1}
\times
\begin{pmatrix} 
\frac{1}{\sigma_{\theta}^2} & 0 & 0 & 0 & ...\\ 
0 & \frac{1}{\sigma_{\theta}^2} & 0 & 0 & ...\\ 
0 & 0 & \frac{1}{\sigma_{\theta}^2} & 0 & ...\\ 
0 & 0 & 0 & \frac{1}{\sigma_{\theta}^2} & ... \\
\vdots & \vdots & \vdots & \vdots & \ddots
\end{pmatrix}\mu_\theta
+
\begin{pmatrix} 
\frac{n\bar{y}}{\sigma^2} & 0 & 0 & 0 & ...\\ 
0 & \frac{n\bar{y}}{\sigma^2} & 0 & 0 & ...\\ 
0 & 0 & \frac{n\bar{y}}{\sigma^2} & 0 & ...\\ 
0 & 0 & 0 & \frac{n\bar{y}}{\sigma^2} & ... \\
\vdots & \vdots & \vdots & \vdots & \ddots
\end{pmatrix}
$$

Posterior Covariance, $Q_\theta^{-1}$:

$$
\begin{pmatrix} 
\frac{1}{\sigma_{\theta}^2} & 0 & 0 & 0 & ...\\ 
0 & \frac{1}{\sigma_{\theta}^2} & 0 & 0 & ...\\ 
0 & 0 & \frac{1}{\sigma_{\theta}^2} & 0 & ...\\ 
0 & 0 & 0 & \frac{1}{\sigma_{\theta}^2} & ... \\
\vdots & \vdots & \vdots & \vdots & \ddots
\end{pmatrix}^{-1}
+
\begin{pmatrix} 
\frac{n}{\sigma^2} & 0 & 0 & 0 & ...\\ 
0 & \frac{n}{\sigma^2} & 0 & 0 & ...\\ 
0 & 0 & \frac{n}{\sigma^2} & 0 & ...\\ 
0 & 0 & 0 & \frac{n}{\sigma^2} & ... \\
\vdots & \vdots & \vdots & \vdots & \ddots
\end{pmatrix}^{-1}
$$

# Problem 3

Solving for $a$ in equation (b) yields:

\[
\begin{aligned}
La & = \ell_\theta \\
a & = \frac{\ell_\theta}{L}
\end{aligned}
\]

Plugging in our result for $a$ into equation (c) yields:

\[
\begin{aligned}
L^T{\theta^*} & = a + e \\
& = \frac{\ell_\theta}{L} + e \\
LL^T{}\theta^* & = \ell_\theta + Le \\
Q_\theta{\theta^*} & = \ell_\theta + Le \\
\theta^* & = \frac{\ell_\theta + Le}{Q_\theta} \\
& = {Q_\theta^{-1}}(\ell_\theta + Le) \\
& = {Q_\theta^{-1}}\ell_\theta + {Q_\theta^{-1}}Le \\
& = {Q_\theta^{-1}}\ell_\theta + N\big[0, (Q_\theta^{-1}L)^TI_d(Q_\theta^{-1}L)\big] \\
& = {Q_\theta^{-1}}\ell_\theta + N\big[0, L^T{Q_\theta^{-1}}^T{Q_\theta^{-1}}L\big] \\
& = {Q_\theta^{-1}}\ell_\theta + N\big[0, \frac{1}{Q_\theta}\big] \\
& = {Q_\theta^{-1}}\ell_\theta + N\big[0, Q_\theta^{-1}\big] \\
& \sim N({Q_\theta^{-1}}\ell_\theta, Q_\theta^{-1})
\end{aligned}
\]

Thus we see that $\theta^* \sim N({Q_\theta^{-1}}\ell_\theta, Q_\theta^{-1})$.

# Code Appendix

```{r, eval = FALSE}

## Load necessary packages

library(coda)
library(asbio)



########## Problem 1 ##########



##### Part i #####



num_draws <- 10^4

mean_c <- 1.013
mean_t <- 1.173

## Draw from the scaled inverse chi-square

n_1 <- 32
df <- n_1 - 1

samp_i_1 <- rinvchisq(num_draws, df, scale=1/df)

## Draw from the normal distribution

samp_i_2 <- rnorm(num_draws, mean = mean_c, sqrt(samp_i_1))



##### Part ii #####

## Draw from the scaled inverse chi-square

n_2 <- 36
df <- n_2 - 1

samp_ii_1 <- rinvchisq(num_draws, df, scale=1/df)

## Draw from the normal distribution

samp_ii_2 <- rnorm(num_draws, mean = mean_t, sqrt(samp_ii_1))



##### Part iii #####

## Generate the histogram of mean differences

hist(samp_ii_2 - samp_i_2,
     xlab = "Mean Difference",
     ylab = "",
     main ="Histogram: Mean Difference")



##### Part iv #####

## Construct the HPD interval of mean differences

interval <- as.vector(HPDinterval(as.mcmc(samp_ii_2 - samp_ii_1), prob = 0.95)[1, 1:2])

## Output the result

cat("HPD Interval: ", "(", interval[1], ",", interval[2], ")")

```