---
title: 'Assignment #2'
author: "Elliot Smith"
date: "8/30/2018"
output: pdf_document
---

# Problem 1

## Part a

Here our goal will be to minimize $a$ to show that $a = E[\theta|y]$ is the unique Bayes estimate of $\theta$:

\[
\begin{aligned}
\frac{d}{da}E[L(a|y)] & = \frac{d}{da}{\int}L(\theta, a)p(\theta|y)d\theta \\
& = \frac{d}{da}{\int}(\theta - a)^2p(\theta|y)d\theta \\
& = -2{\int}(\theta - a)p(\theta|y)d\theta \\
& = -2\bigg[{\int}{\theta}p(\theta|y)d\theta - a{\int}p(\theta|y)d\theta\bigg] \\
& = -2\big[E[\theta|y] - a\big] \\
\end{aligned}
\]

$-2\big[E[\theta|y] - a\big] = 0$ when $a = E[\theta|y]$
\newline
\newline
To prove that it is a unique minimizing statistic, we must look at the second derivative:
\newline
\newline
$\frac{d}{da}(-2\big[E[\theta|y] - a\big]) = 2$
\newline
\newline
As $2 > 0$, this shows that it is a unique minimzing statistic.

## Part b

Here our goal will be to show that for any median value of $a$, the derivative of $L(\theta, a)$ will evaluate to $0$.

\[
\begin{aligned}
\frac{d}{da}\big[E[L(a|y)]\big] & = \frac{d}{da}\bigg[{\int}_{-\infty}^a (a - \theta)p(\theta|y)d\theta + {\int}_{a}^\infty (\theta - a)p(\theta|y)d\theta\bigg] \\
& = {\int}_{-\infty}^a \frac{d}{da}(a - \theta)p(\theta|y)d\theta + {\int}_{a}^\infty \frac{d}{da}(\theta - a)p(\theta|y)d\theta \\
& = {\int}_{-\infty}^a p(\theta|y)d\theta + {\int}_{a}^\infty(-1)p(\theta|y)d\theta \\
& = {\int}_{-\infty}^a p(\theta|y)d\theta - {\int}_{a}^\infty p(\theta|y)d\theta \\
& = \frac{1}{2} - \frac{1}{2} \\
& = 0
\end{aligned}
\]

As a result, it has been shown that any posterior median of $\theta$ is a Bayes estimate of $\theta$.
\newline
\newline
Taking the second derivative we again get a positive number, thus again indicating that it is a minimizing statistic.

## Part c

Here our goal will be to show that for any value of $a$, the derivative of $L(\theta, a)$ will evaluate to $0$ where $k_0$ and $k_1$ are nonnegative numbers.

\[
\begin{aligned}
\frac{d}{da}\big[E[L(a|y)]\big] & = \frac{d}{da}\bigg[{\int}_{-\infty}^a k_1(a - \theta)p(\theta|y)d\theta + {\int}_{a}^\infty k_0(\theta - a)p(\theta|y)d\theta\bigg] \\
& = {\int}_{-\infty}^a \frac{d}{da}k_1(a - \theta)p(\theta|y)d\theta + {\int}_{a}^\infty \frac{d}{da}k_0(\theta - a)p(\theta|y)d\theta \\
& = {\int}_{-\infty}^a {k_1}p(\theta|y)d\theta + {\int}_{a}^\infty(-k_0)p(\theta|y)d\theta \\
& = {\int}_{-\infty}^a {k_1}p(\theta|y)d\theta - {\int}_{a}^\infty {k_0}p(\theta|y)d\theta \\
& = k_1 {\int}_{-\infty}^a p(\theta|y)d\theta - k_0 {\int}_{a}^\infty p(\theta|y)d\theta \\
\end{aligned}
\]

Noting that: $k_0 \int_a^\infty p(\theta|y)d\theta = k_0 - k_0{\int_{-\infty}^a} p(\theta|y)d\theta$

\[
\begin{aligned}
k_1 {\int}_{-\infty}^a p(\theta|y)d\theta - k_0 {\int}_{a}^\infty p(\theta|y)d\theta & = k_1{\int_{-\infty}^a}p(\theta|y)d\theta - \bigg[k_0 - k_0{\int_{-\infty}^a} p(\theta|y)d\theta{\bigg]} \\
& = k_1{\int_{-\infty}^a}p(\theta|y)d\theta + k_0{\int_{-\infty}^a} p(\theta|y)d\theta - k_0 \\
& = (k_1 + k_0){\int_{-\infty}^a}p(\theta|y)d\theta - k_0 \\
\end{aligned}
\]

Now setting ${\int_{-\infty}^a}p(\theta|y)d\theta = \frac{k_0}{k_0 + k_1}$ we get our result that any quantile is a Bayes estimate of $\theta$.
\newline
\newline
Taking the second derivative we again get a positive number, thus again indicating that it is a minimizing statistic.

# Problem 2

n = 20
\newline
Sampling Distribution: $y|\theta \sim Binomial(n = 20, \theta)$
\newline
Prior Distribution: $\theta \sim Beta(\alpha = 2, \beta = 20)$
\newline
Posterior Distribution:
\newline
\[
\begin{aligned}
p(\theta|y) & = \binom{n}{y}\theta^y(1 - \theta)^{n - y} \times \frac{\Gamma(\alpha + \beta)}{\Gamma(\alpha)\Gamma(\beta)}\theta^{\alpha - 1}(1 - \theta)^{\beta - 1} \\
& = \binom{n}{y}\frac{\Gamma(\alpha + \beta)}{\Gamma(\alpha)\Gamma(\beta)}\theta^{y + \alpha - 1}(1 - \theta)^{n - y + \beta - 1} \\
& \propto Beta(y + \alpha, (n - y) + \beta) \\
& = Beta(y + 2, (20 - y) + 20)
\end{aligned}
\]

```{r, echo = FALSE}

### Load in the necessary packages

library(ggplot2)
library(coda)



##### Problem 2 #####



## Set the parameter values

y <- 0
n <- 20
alpha <- 2
beta <- 20

## Set the theta values

theta <- seq(from = 0, to = 1, length.out = 1000)

```

## Part i

$y = 0$
\newline

```{r, echo = FALSE, fig.height = 3}

### Part i ###



## Plot the posterior distribution

post_beta_dens <- dbeta(x = theta, shape1 = y + alpha, shape2 = (n - y) + beta)

ggplot() +
    geom_line(aes(x = theta, y = post_beta_dens), col = "blue") +
    geom_line(aes(x = theta, y = dbeta(x = theta, shape1 = alpha, shape2 = beta)), col = "green") +
    geom_vline(aes(xintercept = (y / n)), col = "red") +
    labs(x = "Theta", y = "Density", title = "Beta Density",
         subtitle = "Blue - Posterior Distribution   /   Green - Prior Distribution   /   Red - MLE")

```

From this graph we can see that while the MLE, $\hat{y} = \frac{y}{n}$ is not properly represented the true distribution due to the fact that there were no positive cases in this hospital, the posterior distribution, via the affect of the prior distribution, is impacting our results enough to render a more accurate result.

## Part ii

```{r, echo = FALSE}

### Part ii ###



## Draw a sample from my posterior distribution

post_beta_rand <- rbeta(n = 10000, shape1 = y + alpha, shape2 = (n - y) + beta)

## Compute the HPD interval

bayes_int <- as.vector(HPDinterval(as.mcmc(post_beta_rand), prob = 0.95)[1, 1:2])

## Calculate the MLE

y_hat <- (y / n)

## Compute the frequentist interval

freq_int <- c((y_hat) - (1.96 * (sqrt(y_hat * (1 - y_hat)) / 2)),
              (y_hat) + (1.96 * (sqrt(y_hat * (1 - y_hat)) / 2)))

## Output the results

cat("HPD Interval: ", "(", bayes_int[1], ",", bayes_int[2], ")")
cat("Frequentist Interval: ", "(", freq_int[1], ",", freq_int[2], ")")

```

From this result, we can see that the Frequestist result compares poorly ($(0, 0)$ is not really an interval at all) against the Bayesian posterior credible interval of $\theta$. Again, the frequentist approach for this hospital is being adversely affected by the fact that there are no positive cases, and as such, we do not get a good representation of the true distribution of positive cases. Whereas the Bayesian approach factors in our prior beliefs on the number of positive cases, thus allowing us to interpret the results for this case.

## Part iii

```{r, echo = FALSE}

### Part iii ###



## Compute the log odds of our posterior beta distrbution

odds_samp <- log(post_beta_rand / (1 - post_beta_rand))

## Compute the HPS interval

bayes_int_odds <- as.vector(HPDinterval(as.mcmc(odds_samp), prob = 0.95)[1, 1:2])

## Output the result

cat("HPD Interval: ", "(", bayes_int_odds[1], ",", bayes_int_odds[2], ")")

```

# Problem 3

## Part i

My methodology for selecting my $\alpha$ and $\beta$ for my $\theta_1, \theta_2 \sim Gamma(\alpha, \beta)$ distribution is as follows. First, using a Pew Research Center on Social & Demographic Trends article on average number of children per parent (http://www.pewsocialtrends.org/2015/05/07/family-size-among-mothers/), I deduced that the average number of children per parent for women in their 20s during the 1970s was 2.28, while the variance is 0.8905. These values were obtained by the results from the poll. Now that I had values for the prior mean and variance, I was able to algebraically solve for $\alpha$ and $\beta$ via the following formulae: $EX = \frac{\alpha}{\beta}$ and $VarX = \frac{\alpha}{\beta^2}$.

```{r, echo = FALSE}

### Part i ###



## Sample mean and variance

samp_mean <- ((1 * 21) + (2 * 43) + (3 * 23) + (4 * 13)) / (21 + 43 + 23 + 13)
samp_var <- round(var(c(rep(1, 21), rep(2, 43), rep(3, 23), rep(4, 13))), 4)

## Compute the gamma parameters

find_gamma_params <- function(mu, sigma_sq) {
    
    beta <- mu / sigma_sq
    alpha <- beta * mu
    return(c(alpha, beta))
    
}

gamma_params <- round(find_gamma_params(samp_mean, samp_var), 4)

cat("Gamma Parameters")
cat("Alpha Parameter: ", round(gamma_params[1], 4))
cat("Beta Parameter: ", round(gamma_params[2], 4))

```

## Part ii

Posterior Distributions:
\newline
\[
\begin{aligned}
p(\theta|y) & = p(\theta_1, \theta_2|y_1, y_2) \\
& \propto p(y_1|\theta_1)p(y_2|\theta_2)p(\theta_1)p(\theta_2) \\
& = \frac{{\theta_1}^{y_1}e^{-\theta_1}}{y_1!}\frac{{\theta_2}^{y_2}e^{-\theta_2}}{y_2!}\frac{\beta^{\alpha}}{\Gamma(\alpha)}{\theta_1}^{\alpha - 1}e^{-{\beta}\theta_1}
\frac{\beta^{\alpha}}{\Gamma(\alpha)}{\theta_2}^{\alpha - 1}e^{-{\beta}\theta_2} \\
& \propto {\theta_1}^{y_1}e^{-\theta_1}{\theta_2}^{y_2}e^{-\theta_2}{\theta_1}^{\alpha - 1}e^{-{\beta}\theta_1}
{\theta_2}^{\alpha - 1}e^{-{\beta}\theta_2} \\
& = \theta_1^{y_1 + \alpha - 1}\theta_2^{y_2 + \alpha - 1}e^{-[{\theta_1}(1 + \beta) + {\theta_2}(1 + \beta)]}
\end{aligned}
\]

Since we may assume $iid$ priors, we get:

$p(\theta_1|y_1) \sim Gamma(y_1 + \alpha, n_1 + \beta)$
\newline
\newline
$p(\theta_2|y_2) \sim Gamma(y_2 + \alpha, n_2 + \beta)$
\newline

```{r, echo = FALSE, fig.height = 3}

### Part ii ###



## Set theta values

theta <- seq(from = 1, to = 5, by = 0.01)

## Set the sample values

y_1 <- 217
y_2 <- 66
n_1 <- 111
n_2 <- 44
lambda_1 <- 217 / 111
lambda_2 <- 66 / 44

## Generate samples from the posterior distributions

post_dens_y1 <- dgamma(x = theta, shape = y_1 + gamma_params[1], rate = gamma_params[2] + n_1)
post_dens_y2 <- dgamma(x = theta, shape = y_2 + gamma_params[1], rate = gamma_params[2] + n_2)
prior_dens <- dgamma(x = theta, shape = gamma_params[1], rate = gamma_params[2])

## Generate the plot

ggplot() +
    geom_line(aes(x = theta, y = post_dens_y1), col = "blue") +
    geom_line(aes(x = theta, y = post_dens_y2), col = "green") +
    geom_line(aes(x = theta, y = prior_dens), col = "orange") +
    geom_vline(aes(xintercept = lambda_1), col = "red") +
    geom_vline(aes(xintercept = lambda_2), col = "purple") +
    labs(x = "Theta", y = "Density", title = "Gamma Density",
         subtitle = "Blue - y_1 Posterior / Green - y_2 Posterior / Red - y_1 MLE / Purple - y_2 MLE / Orange - Prior")

```

As we can see from the plot, we have a fairly non-informative prior; its very hard to tell whether or not our prior distribution is influencing either of our posterior distributions in a meaningful way. Our MLE mean results appear to align very well with our posterior distributions.

## Part iii

```{r, echo = FALSE}

### Part iii ###



## Draw a sample from each posterior distribution

post_gamma_y1 <- rgamma(n = 10000, shape = y_1 + gamma_params[1], rate = n_1 + gamma_params[2])
post_gamma_y2 <- rgamma(n = 10000, shape = y_2 + gamma_params[1], rate = n_2 + gamma_params[2])

## Compute the HPD intervals

hpd_int_y1 <- as.vector(HPDinterval(as.mcmc(post_gamma_y1), prob = 0.95)[1, 1:2])
hpd_int_y2 <- as.vector(HPDinterval(as.mcmc(post_gamma_y2), prob = 0.95)[1, 1:2])

## Output the results

cat("Posterior Credible Interval, theta_1: ", "(", hpd_int_y1[1], ",", hpd_int_y1[2], ")")
cat("Posterior Credible Interval, theta_2: ", "(", hpd_int_y2[1], ",", hpd_int_y2[2], ")")

```

As we can see from our results, the two intervals do overlap!

## Part iv

```{r, echo = FALSE}

### Part iv ###



## Get the difference of our random samples

post_diff_gamma <- post_gamma_y1 - post_gamma_y2

## Compute the HPD interval

diff_int <- as.vector(HPDinterval(as.mcmc(post_diff_gamma), prob = 0.95)[1, 1:2])

## Output the results

cat("Posterior Credible Interval, theta_1 - theta_2: ", "(", diff_int[1], ",", diff_int[2], ")")

```

As we can see from our results, the interval of $\theta_1 - \theta_2$ does include zero!


