---
title: 'Assignment #5'
author: "Elliot Smith"
date: "9/23/2018"
output: pdf_document
---

# Problem 1

```{r, echo = FALSE}

## No scientific notation

options(scipen = 999)

########## Problem 1 ##########



## Input the known parameters and data

y <- c(15,9,15,8,8,20,25,31,13,12,11,15,11,13,13,11,11,12,6,14)
n <- length(y)
S <- 10000
alpha <- 5
beta <- alpha * 15

```

## Part i

```{r, echo = FALSE}

##### Part i #####



## Sample from the Gamma

samp_1i <- rgamma(n = S, shape = n + alpha, rate = beta + sum(y))

```

Prior Distribution: $p(\theta) = \frac{\beta^\alpha}{\Gamma(\alpha)}\theta^{\alpha - 1}e^{-\beta\theta}$
\newline
Likelihood: $p(y|\theta) = {\theta}e^{-\theta{y}}$
\newline
\newline
Posterior Distribution:
\newline
\[
\begin{aligned}
p(\theta|y) & \propto p(y|\theta)p(\theta) \\
& \propto \frac{\beta^\alpha}{\Gamma(\alpha)}\theta^{\alpha - 1}e^{-\beta\theta} \times {\theta}^ne^{-\sum_{i=1}^n\theta{y_i}} \\
& \propto \theta^{(\alpha + n) - 1}e^{-(\beta + \sum_{i=1}^n y_i)\theta} \\
& \sim Gamma(\alpha + n, \beta + \sum_{i=1}^n y_i)
\end{aligned}
\]
\newline
Please refer to the Code Appendix section for the sampling technique used.

## Part ii

```{r, echo = FALSE}

##### Part ii #####



## Construct the replicate data sets

y_rep <- matrix(0, nrow = S, ncol = n)

for(s in 1:S) {
    
    y_rep[s,] = rexp(n = n, rate = samp_1i[s])
    
}

```

Please refer to the Code Appendix section for the technique used to construct the replicate data set.

## Part iii

Please refer to the Code Appendix section for how the test statistics were computed.

```{r, echo = FALSE, fig.height = 6}

##### Part iii #####



## Compute the required test statistics for each replicate data set

y_rep_min <- apply(y_rep, 1, min)
y_rep_q25 <- apply(y_rep, 1, quantile, probs = 0.25)
y_rep_med <- apply(y_rep, 1, median)
y_rep_sd <- apply(y_rep, 1, sd)

par(mfrow = c(2, 2))

hist(x = y_rep_min)
abline(v = min(y), col = 'blue')

hist(x = y_rep_q25)
abline(v = quantile(y, 0.25), col = 'blue')

hist(x = y_rep_med)
abline(v = median(y), col = 'blue')

hist(x = y_rep_sd)
abline(v = sd(y), col = 'blue')

```

## Part iv

```{r, echo = FALSE}

##### Part iv #####



## Compute the posterior predictive p-values for each statistic

pval_min <- mean(y_rep_min > min(y))
pval_q25 <- mean(y_rep_q25 > quantile(y, 0.25))
pval_med <- mean(y_rep_med > median(y))
pval_sd <- mean(y_rep_sd > sd(y))

## Output the results

cat("Minimum p-value: ", round(pval_min, 4))
cat("25% Quantile p-value: ", round(pval_q25, 4))
cat("Median p-value: ", round(pval_med, 4))
cat("Standard Deviation p-value: ", round(pval_sd, 4))

```

Our results imply that our model not be correct since many of our p-values are quite different from our observed value; this tells us that it is very unlikely that the observed value of each statistic lies in our posterior predictive distributions. This suggests that the quantity we computed in class may be incorrect since the median observed values is larger than 10 and our calculated distribution shows a similar result.





